### 2022-01-31 papers/2021-rewriting-popl/response-point-by-point.txt

We use the example of subsection 1.1, where we write:
\begin{minted}[fontsize=\small]{coq}
Make rewriter := Rewriter For (zero_plus, plus_zero, times_zero, times_one, eval_map,
  eval_fold_left, do_again eval_length, do_again eval_combine,
  eval_rect nat, eval_rect list, eval_rect prod) (with delta) (with extra idents (seq)).
\end{minted}

\begin{enumerate}
\item
  The given lemma statements are scraped for which named functions and types the rewriter package will support.
  In this case, the lemma statements are \mintinline{coq}{zero_plus}, \mintinline{coq}{plus_zero}, etc.; the types are \mintinline{coq}{nat}, \mintinline{coq}{list}, and \mintinline{coq}{prod}; and the named functions are \mintinline{coq}{Nat.add}, \mintinline{coq}{Nat.mul}, \mintinline{coq}{List.map}, \mintinline{coq}{List.fold_left}, \mintinline{coq}{List.length}, \mintinline{coq}{List.combine}, \mintinline{coq}{nat_rect}, \mintinline{coq}{list_rect}, \mintinline{coq}{prod_rect}, and the extra identifier given \mintinline{coq}{List.seq}.
\item
  Inductive types enumerating all available primitive types and functions are emitted.
  In this example, the inductive types are approximately
\begin{minted}{coq}
Inductive base := nat | list (_ : base).
Inductive ident := Nat_add | Nat_mul | List_map | List_fold_left | List_length | List_combine | nat_rect | list_rect | prod_rect | List_seq.
\end{minted}
  In fact this list is not accurate, because some types are forced to be included (\mintinline{coq}{nat} and \mintinline{coq}{bool}) while others have hard-coded support separate from the list of base types (\mintinline{coq}{list}, \mintinline{coq}{option}, \mintinline{coq}{unit}), and the list of identifiers is pre-filled with some hard-coded support necessary for making the rewriter work at all.
  The true inductive types generated are:
\begin{minted}{coq}
Inductive base : Set :=  base_nat : base | base_bool : base.
Inductive ident : type (base.type base) → Type :=
| ident_seq : ident (type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat) -> type.base (base.type.list (base.type.type_base base_nat)))
| ident_combine : ∀ t t0 : base.type base, ident (type.base (base.type.list t) -> type.base (base.type.list t0) -> type.base (base.type.list (t * t0)))
| ident_length : ∀ t : base.type base, ident (type.base (base.type.list t) -> type.base (base.type.type_base base_nat))
| ident_fold_left : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0 -> type.base t) -> type.base (base.type.list t0) -> type.base t -> type.base t)
| ident_map : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0) -> type.base (base.type.list t) -> type.base (base.type.list t0))
| ident_succ : ident (type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat))
| ident_mul : ident (type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat))
| ident_O : ident (type.base (base.type.type_base base_nat))
| ident_add : ident (type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat) -> type.base (base.type.type_base base_nat))
| ident_literal : ∀ t : base, match t with | base_nat => nat | base_bool => bool end → ident (type.base (base.type.type_base t))
| ident_nil : ∀ t : base.type base, ident (type.base (base.type.list t))
| ident_cons : ∀ t : base.type base, ident (type.base t -> type.base (base.type.list t) -> type.base (base.type.list t))
| ident_Some : ∀ t : base.type base, ident (type.base t -> type.base (base.type.option t))
| ident_None : ∀ t : base.type base, ident (type.base (base.type.option t))
| ident_pair : ∀ t t0 : base.type base, ident (type.base t -> type.base t0 -> type.base (t * t0))
| ident_tt : ident (type.base ())
| ident_prod_rect_nodep : ∀ t t0 t1 : base.type base, ident ((type.base t -> type.base t0 -> type.base t1) -> type.base (t * t0) -> type.base t1)
| ident_bool_rect : ∀ t : base.type base, ident ((type.base () -> type.base t) -> (type.base () -> type.base t) -> type.base (base.type.type_base base_bool) -> type.base t)
| ident_list_case : ∀ t t0 : base.type base, ident ((type.base () -> type.base t0) -> (type.base t -> type.base (base.type.list t) -> type.base t0) -> type.base (base.type.list t) -> type.base t0)
| ident_option_rect : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0) -> (type.base () -> type.base t0) -> type.base (base.type.option t) -> type.base t0)
| ident_nat_rect : ∀ t : base.type base, ident ((type.base () -> type.base t) -> (type.base (base.type.type_base base_nat) -> type.base t -> type.base t) -> type.base (base.type.type_base base_nat) -> type.base t)
| ident_eager_nat_rect : ∀ t : base.type base, ident ((type.base () -> type.base t) -> (type.base (base.type.type_base base_nat) -> type.base t -> type.base t) -> type.base (base.type.type_base base_nat) -> type.base t)
| ident_nat_rect_arrow : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0) -> (type.base (base.type.type_base base_nat) -> (type.base t -> type.base t0) -> type.base t -> type.base t0) -> type.base (base.type.type_base base_nat) -> type.base t -> type.base t0)
| ident_eager_nat_rect_arrow : ∀ t t0 : base.type base, ident ((type.base t -> type.base t0) -> (type.base (base.type.type_base base_nat) -> (type.base t -> type.base t0) -> type.base t -> type.base t0) -> type.base (base.type.type_base base_nat) -> type.base t -> type.base t0)
| ident_list_rect : ∀ t t0 : base.type base, ident ((type.base () -> type.base t0) -> (type.base t -> type.base (base.type.list t) -> type.base t0 -> type.base t0) -> type.base (base.type.list t) -> type.base t0)
| ident_eager_list_rect : ∀ t t0 : base.type base, ident ((type.base () -> type.base t0) -> (type.base t -> type.base (base.type.list t) -> type.base t0 -> type.base t0) -> type.base (base.type.list t) -> type.base t0)
| ident_list_rect_arrow : ∀ t t0 t1 : base.type base, ident ((type.base t0 -> type.base t1) -> (type.base t -> type.base (base.type.list t) -> (type.base t0 -> type.base t1) -> type.base t0 -> type.base t1) -> type.base (base.type.list t) -> type.base t0 -> type.base t1)
| ident_eager_list_rect_arrow : ∀ t t0 t1 : base.type base, ident ((type.base t0 -> type.base t1) -> (type.base t -> type.base (base.type.list t) -> (type.base t0 -> type.base t1) -> type.base t0 -> type.base t1) -> type.base (base.type.list t) -> type.base t0 -> type.base t1)
| ident_List_nth_default : ∀ t : base.type base, ident (type.base t -> type.base (base.type.list t) -> type.base (base.type.type_base base_nat) -> type.base t)
| ident_eager_List_nth_default : ∀ t : base.type base, ident (type.base t -> type.base (base.type.list t) -> type.base (base.type.type_base base_nat) -> type.base t)
.
\end{minted}
  We believe that this is long enough as to be confusing rather than illuminating.
\item
  Tactics generate all of the necessary definitions and prove all of the necessary lemmas for dealing with this particular set of inductive codes.
  Definitions include operations like Boolean equality on type codes and lemmas like ``all representable primitive types have decidable equality.''
  There are actually 69 definitions and lemmas generated during this step, and it's needlessly confusing to list all of them.
  They can be found by inspecting the records \mintinline{coq}{Rewriter.Language.IdentifiersBasicLibrary.Compilers.Basic.GoalType.package} and \mintinline{coq}{Rewriter.Language.IdentifiersLibrary.Compilers.pattern.ident.GoalType.package}.
  Note that the first of these has a couple of fields which are themselves records, and the definition of these records must also be inspected for the full picture.
\item
  The statements of rewrite rules are reified and soundness and syntactic-well-formedness lemmas are proven about each of them.
  Each instance of the former involves wrapping the user-provided proof with the right adapter to apply to the reified version.
  It is not clear that examples would help here.
\item
  The definitions needed to perform reification and rewriting and the lemmas needed to prove correctness are assembled into a single package that can be passed by name to the rewriting tactic.
  This step is semantically trivial but extremely verbose.
\end{enumerate}

Overall, it's not clear what this reviewer was hoping to get from a
worked out example.

> - The micro-benchmark 4(a) does not include the reification time, which seems a little strange given that reification is very much part of the technique.

The micro-benchmark 4(a) includes lines both with and without
reification.  The former is labeled "Our approach including
reification, cbv, etc." while the latter is labeled "Our approach
(only rewriting)".  The lines are quite close, and both are
significantly better than other approaches.  Moreover, as explained in
the benchmark of 4(c) (the only benchmark where the time of just
rewriting diverges significantly from the full time), more than 99% of
the overhead is due to *interpretation* in Coq's cbv tactic.

In all microbenchmarks, we have constructed the input so that it is
trivial to reify, and the overhead is negligable when there is any
real rewriting work to be done.  Reification time in the
macrobenchmark is significant due to the fact that we reify modulo δ
without sharing of subterms, but this is an artifact of the way that
we've integrated the rewriter into Fiat-Crypto.

Finally, as we explain at the end of the first paragraph of 5.1.1 on
lines 785--788, the particular method of reification is not integral
to the framework, and we have prior work demonstrating how to achieve
extremely fast reification.

> - Does the technique allow for compilation of P-384 in Fiat Crypto? The introduction mentions it, but the experiments never come back to it.

Yes.  P-384 is a prime close to 2^384, and figure 4(d) shows that our
benchmarks extend well past 2^400.  In fact, though, as described in
section 5.2, "in the course of running this experiment, we found a way
to improve the old approach for a fairer comparison."  This
improvement allows the old approach to compile P-384 (32-bit) in about
55 seconds.  We achieve about 16 seconds if you extract and compile
our code, or about 117 seconds if you run our code in Coq's VM.

Responses to Review #81C
===========================================================================

> That said, I doubt that even an expert could reproduce what they have
> done, based on this paper.  There's an overview in 2.1, aspects of
> which are fleshed out later, but a great deal is unspecified.  In a
> paper of this length that's probably inevitable.  But it's still sad.

This is true.  We have another 28 pages of content merely describing
the trickiest implementation and design choices, and the document
describing the full design of the core of the framework (not including
the code that makes the rewriter convenient to use, i.e., only
including the rewriter itself and the correctness theorem used in step
(3) of the second half of our nine-step approach in section 2.1) is a
bit over 40 pages.  Note that this document does not describe any of
the proof techniques, only the definitions and proof statements.
Documenting the full rewriter is undoubtedly a much larger effort, and
would not fit in a paper such as this.

> It's a description of a substantial and sophisticated engineering
> achivement, not a crisp, pearly idea.

Indeed, and this is unfortunate.  If there is a crisp, pearly idea to
be found, beyond assembling preexisting techniques, it is that in the
tradeoff between slow generic procedures and fast domain-specific
code, we can have our pie and eat it too by specializing the generic
reflective method to our specific domain and partially evaluating it
in a pre-compilation phase.  However, this paper is not about this
pearl, but about the larger system in which it is applied.

Responses to Comments for author
-------------------
Some small notes

> * Fiat Crypto.  Line 897 says you replicate Erbsen; and yet on line 55
>   you say you can do something they can't.  Which is right?  If they
>   can't do P-384, how is it done in deployed systems?  Do the deployed
>   systems use verified code?  If not, how does the verified code
>   compared with the (presumably hand-written) deployed code?  Please
>   be clearer about all this.

We say on lines 910--911, "[a]ctually, in the course of running this
experiment, we found a way to improve the old approach for a fairer
comparison."  This modification of their approach allows synthesis of
P-384.

Perhaps the following modifications would make the text clearer?  To
the first sentence ("[...] replicating the generation of
performance-competitive finite-field-arithmetic code for all popular
elliptic curves by Erbsen et al. [2019]"), there is an implicit "as
well as generating code for a number of curves they were unable to
generate code for".  The second sentence, "In all cases, we generate
essentially the same code as they did," should instead read "In all
cases where they succeeded in generating code, [...]".

The point of this comparison is that we are generating code from the
same starting point and with the same ending point, but we can handle
more and larger primes than could be handled in Erbsen et al. [2019].

> * Line 93.  Special marker.  Is this part of Coq already, or are you
>   inventing it?
>
> * Line 96 "Further marker".  Same question.

Both of these are just the identity function under a different name.  That is, we write
```
Definition literal {T} (v : T) := v.
Notation "' x" := (literal x).
```
for the apostrophe marker and we write
```
Definition eagerly {T} (v : T) := v.
```
for the further marker.
```

> * Line 106. I'm already.  I'm not a Coq user so I don't know
>   what plugins do.  Is do_again defined by you or built-in?  "request
>   that some rules trigger a bottom-up pass.."  Who implements that
>   request.  It's all very baffling.

Appologies for the confusion.

A Coq plugin allows us to extend not just the tactic language but the
vernacular language with new commands; we use it to create a single
command that the user can type which will define the needed
definitions and inductive types.

When we speak of "requests" and "triggers", we are speaking of
annotations which the user gives and which our rewrite procedure
inspects to determine what to do.  As a very simple example, you could
write a tactic which takes in a pair of a boolean and a lemma, and
rewrites with the lemma once when the boolean is false, but rewrites
repeatedly when the boolean is true.  This procedure could be written
in Ltac as
```
Ltac my_procedure argument :=
  lazymatch argument with
  | (true, ?lemma) => repeat rewrite lemma
  | (false, ?lemma) => rewrite lemma
  end.
```
You could say that passing the true boolean is a request to the
procedure which triggers repeated rewriting, and this is in fact
almost exactly how our rewriter functions.

To answer the last part of this question in full detail, we define
`do_again` as another identity function:
```
Notation do_again := (@id (@snd bool Prop (pair true _))) (only parsing).
```
(Note the use of `pair true`, which is us annotating the lemma with a
boolean indicating that rewriting should be repeated.  We perform the
annotation as an argument to the identity function merely for
convenience of recognition by our tactic.)

> * Line 226 little examples like this, illustrating what the perf
>   challenges are, are really helpful.  More please.

Glad to hear this!  We like small examples like this too, but they
tend to be notoriously hard to come by, especially in Coq and other
proof assistants where there is a great deal going on under the hood,
and it's not always clear where the bottlenecks are.  We'd be
interested in knowing if there are specific other parts of the paper
that could benefit from examples like these.

### 2022-01-31 papers/2022-rewriting-popl/response.txt

Thanks for the reviews, and we'll address the seemingly largest concerns before returning to answer additional questions.

The biggest concern seems to be about generality of the approach.
There is enough going on in the Fiat Cryptography example (almost 200 distinct rewrite rules covering quite a range of functional-programming patterns) that quantifying it as just "one compiler" may be misleading.
Still, we acknowledge that, beyond the microbenchmarks, we've only assessed this approach as an improvement to Fiat Cryptography.
However, our tool is a fully free-standing Coq plugin, and making use of it requires only a couple lines of code in the simplest cases (see Appendix C in the supplement for details).
We're pretty confident that we can get any Coq expert started using it on their own intuitive examples in 10 minutes; a sanity check of this kind could even be reasonable as a required revision.

We expect that our approach would apply smoothly to any compiler which can be phrased in terms of rewrite rules.
Our particular tool is subject to some minor engineering limitations, such as requiring that the source, target, and all intermediate languages be simply typed, and some less-minor limitations, such as, at present, requiring that the source and target language be total and that the semantics be expressed in terms of a denotation function.
The reviewers are probably right that our approach applies more naturally to domain-specific languages/compilers (e.g., we don't save much effort when it comes to program-analysis phases), but we are interested to discuss with authors of the cited compiler-verification projects about opportunities to streamline with our tooling.

Additionally, two reviewers asked about the rewriting strategy we implement and how this strategy is controlled by the user.

Our rewriting strategy, described in detail in section 3.3, is normalization-by-evaluation-order rewriting.
This is closest to a bottom-up strategy, except that it has much better behavior on lambda-abstractions, achieving asymptotically better performance than either top-down or bottom-up rewriting in some cases.
We now plan to add a paragraph calling out this strategy more explicitly, emphasizing that we believe it's the correct one for rewriting higher-order terms.
Rewrite rules are then performed in-order; more granular phasing of rewrite rules can be performed by sequencing invocations of the rewriter.
As described in section 2, users can specify that certain rewrite rules should trigger additional rewriting in their output (with `do_again`), allowing for a limited form of top-down rewriting.

As an example, consider rewriting with the rule `map f xs = list_rect _ [] (fun x _ map_tl => f x :: map_tl) xs` together with the rule that says to evaluate list recursion, `list_rect`, on concrete lists.
If we want to rewrite in an expression such as `map succ (map succ [0; 1; 2; 3])`, we must mark the `map` rule as `do_again` to achieve complete rewriting in a single pass.

# More responses past 500 words (other important high-level points)

Reviewer B asked:
> What about conditional rewriting rules in the style of OCaml's "when" or Agda's "with abstractions"?

The side conditions described in section 4.3 support the full scope of OCaml's "when" and Agda's "with abstractions".
The only restriction is that conditions can only be applied to fully evaluated terms (we cannot, for example, say that `x + y` should be rewritten into `2 * x` when `x == y` if `x` and `y` are not known to be fully evaluated).
Users specify side conditions for our engine by implication, so OCaml's `| pat when cond -> body` becomes `cond = true -> pat = body`, and Agda's `f args with term` followed by `f pat1 | pat2 = body` becomes, roughly `forall args, term = pat2 -> pat1 = body`.

Reviewer C asked:
> A claimed advantage over Aehlig et al. is that the trusted computing base is not grown.
> However, Aehlig et al. claim (page 2 of their paper) that their approach does not grow the trusted base either, "if the compiler used at runtime is the same as the one compiling the theorem prover".
> Could you respond to this claim?

As we understand it, Aehlig et al. formalize a semantics of ML, code up their rewriting engine relative to that semantics of ML, and then use a (trusted, unverified) compiler for ML to compile their rewrite engine and rewrite rules.
In section 3 (Model and Verification), they say "we over-approximate the operational semantics of ML."
This model that is an over-approximation of ML is an extension of the TCB, unless the ML compiler is proven with respect to this model.
(Granted, using a different compiler for rewriting and for the proof assistant would grow the TCB much more.)

Reviewer D asked:
> Could you explain how the compiler works (and is used) when it is extracted to OCaml code?
> In this scenario, what source language and target language does it use?

Fundamentally, the extracted compiler consumes and produces ASTs of the kind introduced in this paper, for a subset of Gallina.
Building a full real-world compiler requires statically composing the extracted heart of the compiler with two additional pieces: a way to input or generate deeply-embedded syntax, and a way to print or consume the resulting deeply-embedded output.
In Fiat Cryptography, we reify once and for all a template library of arithmetic routines, extracting the deeply embedded syntax trees in addition to the compiler.
We additionally have (mostly unverified) backends that pretty-print the compiled AST to strings, which we also extract to OCaml.
This results in a binary that can parse elliptic curve parameters from the command line, apply pre-reified ASTs to these parameters, run the rewriter to produce compiled code, and pretty-print the resulting ASTs to stdout.
Hence the source language of the compiler is Gallina (though in some sense the source language of our extracted binary can be said to be a tiny DSL of elliptic curve parameters), and the target language of the compiler is ASTs, which are then processed to emit C, Java, Rust, or Go code.
In theory, we could, for example, write a translator from our ASTs to CompCert C and then compose our pipeline with the CompCert compiler, which would allow us to have extracted binaries which produce object files from elliptic curve parameters.

# Answering remaining questions per review

## Review A

> line 232: This sentence seems to be missing a word.

Perhaps the subject change in "That compiler execution ran inside of Coq, but even more pragmatic is to *extract* the compiler as a standalone program in OCaml or Haskell." is confusing, and we should have instead written "That compiler execution ran inside of Coq, but an even more pragmatic approach is to *extract* the compiler as a standalone program in OCaml or Haskell."

> line 749: You refer to the developers of Fiat Crypto as 'we'; this is inconsistent with other places in the paper.

Oops, thanks.
I think for the camera-ready submission we will probably be changing all instances of "The Fiat Cryptography authors" to "we".

> line 798: Please define IR (I suppose it is 'intermediate representation')

Yes, that's correct.  We'll fix this.

## Review B

> Section 3.3: the motivating example could be treated by first beta-reducing to $z + 0$, then rewriting as usual.
> I thought you wanted to demonstrate rewriting under lambdas?
> In which case, $\lambda x. x + 0$ would be a better example.

In fact we achieve something much better than just rewriting under lambdas: we are able to interleave reduction and rewriting in a way that only a single pass over the term is required for full reduction, in most cases, even when rewrite rules expose new beta redexes and even when beta reduction exposes new rewriting opportunities.
A more complete example might have been $\texttt{map }(\lambda x. x + 1)\ [0, x, y]$, where rewriting exposes the $\beta$-redexes $[(\lambda x. x + 1)\ 0,\ (\lambda x. x + 1)\ 1,\ (\lambda x. x + 1)\ 2]$, $\beta$-reduction exposes rewriting opportunities in $[0 + 1, x + 1, y + 1]$, and finally rewriting reduces this to $[1, x + 1, y + 1]$.
We were concerned that this might be too complicated or nuanced of an example just for orienting the reader to NbE.

> Section 4.2: yes, subterm sharing is crucial, but sometimes terms must be unshared so that rewriting / reduction can proceed.
> (That's the optimal reduction dilemma.)
> Strategies such as "let-reduce when the bound term is a constant" are simplistic.
> There are many other heuristics in the literature on compiling functional languages, e.g. Appel's book "Compiling with continuations", but maybe what is needed is ways for users to control "let" reductions?

That is a fair point, which we have indeed avoided confronting because we are exploring a rather different compilation style than is most standard for functional languages, thanks to the opportunity to reduce away all higher-order features at compile time.
It is still true that any pattern that can be expressed as a rewrite rule with Boolean side conditions can be used for inlining "let"-binders, and we also support some additional heuristics.

> Section 4.4 and section 5.1.4: it is really unclear how bit-width of parameters and intermediate results is being tracked.
> I would have expected types "bits(N)" to be used for this purpose, with specific passes to infer appropriate bit widths, but this is not what happens, apparently.

Bitwidths are tracked using the arguments to the $\text{clip}_{l,u}$ functions in section 4.4.
There is a specific pass, outside the rewriting engine, to infer these bitwidths via abstract interpretation (though, as mentioned in 5.1.6, having abstract interpretation outside the rewriting engine results in quartic performance where we ought to be able to achieve quadratic performance).
We could have used "bits(N)" types, as long as we hard-coded the permissible values of N ahead of time and defined separate versions of each function for each value of N, because our engine doesn't yet support dependent types.
However, we saw no benefit to using bits(N) types over the more general, less-dependently typed solution of using `clip` functions.

## Review C

> I was quite confused by the third paragraph of the intro.
> The paragraph is explaining the scalability challenges of _proof-producing_ compilers.
> But the next paragraph goes on to say that this paper is about a _proven-correct_ compiler, which doesn't have these challenges.
> So I didn't understand why that third paragraph is necessary.
> The reader has already been primed to expect proven-correct compilers from the very first line of the introduction, so can't proof-producing compilers be deferred to Related Work?

We believe this paragraph is important for motivating our work.
Proof-producing is almost always simpler, easier to manage, easier to debug, and easier to maintain than proven-correct.
Hickey and Nogin's [2006] work, when cast in the frame of modern proof assistants such as Coq, Agda, or Isabelle/HOL, is about proof-producing compilers.
We originally tried to have Fiat Cryptography be a proof-producing compiler (see Appendix D).
Scalability challenges of proof-producing compilers are what motivated us to develop a toolkit for building proven-correct compilers in the first place, and performance concerns are essentially the only reason we see to maintain a proven-correct compiler for Fiat Cryptography rather than a proof-producing one.
Since performance concerns are so front-and-center in motivating our particular tool, we believe this paragraph deserves its place in the introduction.

> I was a bit surprised that Hickey and Nogin's [2006] work on compilers built using rewrite rules is criticised for its rewriting engine being _unproven_.
> As I understand Hickey and Nogin's work, that's kinda the point -- once the rewrites are proven, it doesn't matter which order they're applied in, so the rewriting engine can be unproven (and hence easier to change, etc).
> But your point about them not supporting side-conditions seems believable.

As we understand it, Hickey and Nogin's claim is not that the rewriting engine can be unproven, but that it can be divided into unproven heuristic parts (determining the order of rule application, performing pattern matching) and a trusted Logical Framework for scope preservation, binder tracking, etc.
Insofar as the Logical Framework maintains any kind of proof object (whether abstract as in Isabelle/HOL or concrete as in Coq), such a rewriting engine is inherently proof-producing rather than proven-correct, and hence subject to hard-to-avoid asymptotic issues in the presence of deeply nested binders.
Indeed, in our framework, order selection and term decomposition for pattern matching are unproven; we only need to prove the part of the framework that composes the rewrite rules; tracks scope preservation, binders, and variable instantiation; and performs reduction.

> Line 312: I'm not sure what "nearly simply typed" means.

"Nearly simply typed" means that we support prenex type-polymorphism in constants so long as the types are always concretely instantiated whenever the constant actually appears in an expression.

## Review D

> My description is probably inaccurate, reflecting my limited understanding of the details.

Your description is accurate.

> The presentation is [...] not self-contained enough, not gentle enough.
> The technical presentation is incomplete: reification is not described at all, I think; the use of decision trees is only briefly described; the treatment of `LetIn` constructs in the NbE algorithm is barely described; etc.

We tried to strike a balance between being overly technical and inaccessible to all but the most expert readers on the one hand, and sweeping so many details under the rug that readers would come away with an inaccurate picture on the other.
To give a sense of scale, we have another thirty pages of written content on the various design decisions, tradeoffs, and lessons of the core of the framework, and approximately another sixty pages sketching out the design of the rewriting engine in full detail.
We aimed to give an overview of the novel features of the rewriting engine, describe the lessons we learned, and present the engine in enough detail that a motivated reader could reconstruct most of the key components.
By and large, we chose to omit descriptions of techniques where there is already a rich existing literature (our engine is compatible with any method of reification, and the particular method we chose is unimportant), where we make no innovations over the existing literature (our implemention of decision trees is a transcription to Coq of the algorithm described in Maranget's "Compiling Pattern Matching to Good Decision Trees"), or where the treatment is straightforward and without pitfalls (the `LetIn` constructs follow the obvious combination of application and abstraction in the NbE algorithm of `reduce(let v := e_1 in e_2) = let x := reduce(e_1) in reduce([x/v]e_2)`).
In Appendix E, we give a detailed map between the description in the paper and the code in the code supplement, which we hope will fill in the gaps which we had no space to cover in the paper.

> lines 29-31, "most desired iteration on the compiler can be achieved through iteration on the rewrite rules".
> The authors seem to suggest that all changes to the compiler can take the form of new rewrite rules, or modifications to existing rewrite rules.
> However, much of the "foresight" that a compiler writer must display resides in the designing of appropriate intermediate languages, separated by conceptual gaps that must be just large enough to be interesting and small enough to be tractable; and in the ordering of these intermediate languages.
> One might argue that, once the design of these ILs has been frozen, some optimisations may become difficult to express, as they might require extending the definition of one or more ILs, or (worse) they might require revisiting the order in which certain transformations take place.

As we discuss in 5.1.2, 5.1.3, and especially 5.1.4, design of the appropriate intermediate languages falls within the scope of adding or tweaking rewrite rules in our framework.
The plugin code will automatically generate identifiers for any definitions mentioned in any rewrite rule.
Users are free to change which definitions they use for a shallowly-embedded IR; no freezing of design is necessary.

> line 33, "Proof-producing compilers usually operate on the functional languages of the proof assistants that they are written in".
> Perhaps I am missing something, but I don't see why that is or must be the case.
> In fact, there is probably a spectrum of "proof-producing compilers", where the "proof" takes the form of a witness that can be checked by a certified verifier; the more complex the verifier, the closer we are to a "certified" compiler, and the simpler the verifier, the closer we are to a "certifying" compiler.
> When the authors write "Proof-producing compilers usually operate [...]", which end of the spectrum do they have in mind?
> (Perhaps the extreme where there is no user-defined verifier and the witness must be a term of type Prop?)

Yes, we are considering the end where the type of a witness has type Prop.
Propositions are about terms in the functional language of the proof assistant.
While it's possible to add extra indirection (as is done, for example, in Iris' proof mode), the simplest solution is to just operate directly on such terms.

> line 60, "Their execution timed out": the execution of what?

Execution of the old Fiat Cryptography framework / code generator.
(We let it run for over 300 hours; it didn't finish.)

> line 90, "preserving subterm sharing".
> It is not clear what kind of sharing the authors have in mind.
> Sharing of program terms, or sharing of proof terms?
> Why are there important opportunities for sharing?
> Why cannot R_tac preserve sharing?

Sharing of program terms.
R_tac does not have native support for subterm sharing, and without the let-lifting optimization described in 4.2, rewriting requires asymptotically more term traversals.

> line 135, "we expect it to be performance-prohibitive to perform bookkeeping operations on first-order-encoded terms".
> Why is that?
> Is it because the authors assume that de Bruijn indices are encoded as unary numbers?
> Or would the cost be prohibitive even with an efficient representation of de Bruijn indices (either as natural integers in binary notation or as machine integers)?

Merely having an efficient representations of de Bruijn indices is insufficient; subsitution is a very significant cost.
In Figure 3(c), the gap between the blue line (our method with the final conversion from a deeply-embedded AST to a shallowly-embedded term) and the green line (just the rewriting and reduction portion of our method) is entirely due to an inefficiency in Coq's handling of de Bruijn indices, which are nevertheless encoded as machine integers.
More details are available in [Coq Issue #11151](https://github.com/coq/coq/issues/11151#issuecomment-557435656), where the foremost expert on Coq performance says:

> My understanding so far is that the cbv / lazy reduction machine are completely broken performance-wise on open terms.
> The explicit substitutions for the CBV machine for instance seem to be enforcing a call-by-name behaviour for lifts, as every time we access a variable we have to crawl the substitution and recompute the lifted result.
>
> Compare this to the VM which uses an accumulator, where closures are represented by mere arrays.
> I think that the NbE approach is inherently more efficient because object-language values are indeed represented by meta-language values.
>
> If anything, this is probably a theoretical issue that requires more pondering.

The fix for this [required using "a variant of skewed lists enriched over a monoid" to track substitutions](https://github.com/coq/coq/pull/13537).
We are happy to defer issues involving efficient handling of binders to the Coq kernel.

One point that this reviewer might be missing is that we are dealing with enormous terms, and that many of the scaling challenges that arise in Fiat Cryptography are not present in any other project we've yet encountered.

> line 142, "Fiat Cryptography's arithmetic routines rely on significant sharing".
> It is still not clear what is meant.
> If sharing matters, why not explicitly represent it via an object-level `let` construct?
> Does the difficulty stem from the fact that a shallow embedding is used, therefore it is difficult to control whether and when Coq decides to unfold its meta-level `let` construct?

We do explicitly represent sharing via an explicit object-level `let` construct.
The issue is that this construct blocks reduction unless either the code is rewritten in CPS or the reduction engine has native support for lifting `let`-binders, and that attempting to support `let`-lifting via rewrite rules results in poor asymptotic performance.
Again, the primary point here is about performance and scaling challenges.

> line 178, I suppose I could have understood this earlier, but it is becoming clear only now that the paper is about partial evaluation, and about a form of supercompilation.

Yes, perhaps we should have brought in the role of partial evaluation earlier.

> line 203, the role of `ident.eagerly` is unclear.
> Is it an identity function?
> I assume that it serves as a marker.
> It would be good to point precisely to the place in the paper where the recognition of this marker is explained.
> Why is `ident.eagerly` *not* used in the lemmas `eval_combine` and `eval_length`?

Yes, it is an identity function that serves as a marker.
Recognition of this marker is hard-coded into the reification of replacement patterns of rewrite rules into NbE-typed values, where these identifiers are replaced by custom definitions which await concrete terms to recurse on.
Appendices E.4.5 and E.2 describe some of the code associated to this feature.
We decided that this detail was too technical and required too much contextual explanation to include in this paper.
The `ident.eagerly` marker could be used in `eval_length`; there is no reason not to do so, and we would then not need to mark `eval_length` with `do_again`.
However, while we could mark the `list_rect` in `eval_combine` with `ident.eagerly`, we would still need `do_again` to handle `list_case`, because we did not build in special support for "eager" evaluation of `list_case`.
(Such support is unnecessary, because the two rewrite rules for the `cons` and `nil` cases for `list_case` are complete with only a single pass of rewriting, while no rewrite rule reified from a Gallina expression can express complete reduction of `list_rect` under a single pass of rewriting, or even under a fixed number of rewriting passes.)
